---
title: "Property Value Prediction Model for Seattle"
author: "Viva/Yaohan"
date: "`r Sys.Date()`"
output:
   html_document:
    keep_md: yes
    toc: yes
    theme: flatly
    toc_float: yes
    code_folding: hide
    number_sections: no
    fontsize:12pt
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = TRUE
  )
options(scipen = 999)

if(!require(pacman)){install.packages("pacman"); library(pacman)}
p_load(tidyverse, sf, units, nngeo, tmap, tinytex, kableExtra, janitor, classInt,
       patchwork, here, tidycensus, lwgeom, lsr, haven, descr, RColorBrewer, stargazer,
       car, FNN, GGally, MASS, ISLR2, spdep, caret, ckanr, grid, gridExtra, ggcorrplot,
       jtools, broom, tufte, rmarkdown)

library(conflicted)
conflicts_prefer(dplyr::select)
conflicts_prefer(dplyr::filter)

source("https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/functions.r")
```


# Introduction

This report elaborate on the process of developing and refining a comprehensive predictive model that accurately forecasts housing prices for King County, Washington (Seattle).

???by integrating both structural and locational attributes. The model aims to discern how various housing characteristics, such as size, age, and amenities, along with locational factors, including proximity to services, neighborhood quality, and accessibility, collectively impact property values.


# Data Wrangling

## Data Selection

**1. Internal Characteristics**
- Year used (continuous)
- Renovation (dummy)
- Bedroom (continuous + category)
- Bathroom (continuous + category)
- Floors (continuous + category)
- Living area (continuous)
- Lot area (continuous)
- Waterfront (dummy)
- View (category)
- Condition (category)
- Grade (category)

**2. Socio-economic Characteristics**
- Population density (continuous)
- White population share (continuous)
- Age structure (continuous)
- Education level (continuous + dummy)
- Median household income (continuous + dummy)
- Employment rate (continuous + dummy)
- Poverty rate (continuous + dummy)

**3. Amenities Services**
- Subway Station (distance + category)
- School District (category)
- Park (area + count)
- Medical Facilities (distance + category)
- Commercial center (distance + count)
- Crime rate (distance + )

reference: https://www.rentseattle.com/blog/how-local-amenities-help-seattle-investors-find-good-properties

**4. Spatial Structure**
- Large district
- Small district

## Boundary and House

```{r Seattle_boundary}
# read Seattle boundary
seattle <- st_read(here::here("data/raw/Boundary/Seattle_City.geojson")) %>%
  st_union() %>%
  st_as_sf() 

ggplot() + 
    geom_sf(data = seattle, color = "#d7191c", linewidth = 0.4) +  # add Seattle boundary
    labs(title = "Seattle, King County") +  # set plot title
    theme_void() +
    theme(plot.title = element_text(hjust = 0.5, face = "bold"))
```

```{r house_location}
# import house data
kc_hh<- read_csv(here::here("data/raw/kc_house_data.csv"))

# quick check of house data
str(kc_hh)
anyNA(kc_hh$lat)
anyNA(kc_hh$long)

# create house location
kc_hh <- kc_hh %>%
  st_as_sf(coords = c("long", "lat"), crs = 4326, agr = "constant") %>%  # convert to sf object with specified CRS
  st_transform(st_crs(seattle)) %>%  # transform coordinate reference system to "Washington State Plane North"
  distinct()  # keep only distinct geometries

hh <- kc_hh %>%
  st_intersection(seattle)

# plot Seattle house location
ggplot() + 
    geom_sf(data = seattle, color = "#b2182b", linewidth = 0.5) +  # add Seattle boundary
    geom_sf(data = hh, color = "#2166ac", size = 0.5, pch = 16) +  # add house location
    labs(title = "Seattle House Location") +  # set plot title
    theme_void() +
    theme(plot.title = element_text(hjust = 0.5, face = "bold"))

#examine house id
length(unique(hh$id))#6691 < 6740
#add an unique key
hh$key <- 1:nrow(hh)

#exclude outliers with extremely high price
hist(hh$price)
hh[hh$price > 3000000,]
hh <- hh %>%
  filter(!price > 5000000)
```

## Internal Characteristics

```{r internal, message = FALSE, warning = FALSE}
summary(hh)

# add year_used and renovation and ensure categorical data are factor
house <- hh %>% 
  mutate(year_used = 2015 - yr_built, # used year
         reno_dum = as.factor(if_else(yr_renovated>0, 1, 0)),#renovation yes or no
         water_dum = as.factor(waterfront),
         view_cat = as.factor(view),
         condition_cat = as.factor(condition),
         grade_cat = as.factor(grade))

# bed categories
house$bed.factor <- factor(house$bedrooms, levels =sort(unique(house$bedrooms)))
house %>%
  st_drop_geometry() %>%
  group_by(bed.factor)%>%
  summarize(price_m = mean(price))%>%
  ggplot(aes(x = bed.factor, y = price_m)) +
  geom_col(position = "dodge")+
  plotTheme() + theme(axis.text.x = element_text(angle = 45, hjust = 1)) #0-3,4-7,8+
# group by mean
house <- house %>%
  mutate(bed_cat = factor(case_when(
    bedrooms <=3 ~ "few",
    bedrooms >3 & bedrooms <= 7 ~ "medium",
    bedrooms >=8 ~ "many"
  )))%>%
  select(-bed.factor)
# check the numbers of each category
table(house$bed_cat)

# bathroom category
house$bath.factor <- factor(house$bathrooms, levels =sort(unique(house$bathrooms)))
house %>%
  st_drop_geometry() %>%
  group_by(bath.factor)%>%
  summarize(price_m = mean(price))%>%
  ggplot(aes(x = bath.factor, y = price_m)) +
  geom_col(position = "dodge")+
  plotTheme() + theme(axis.text.x = element_text(angle = 45, hjust = 1)) #0-4, 4+
house <- house %>%
  mutate(bath_dum = factor(case_when(
    bathrooms <= 4 ~ "few",
    bathrooms > 4 ~ "many"
  )))%>%
  select(-bath.factor)
# check the numbers of each category
table(house$bath_dum)

# floor category
house$floor.factor <- factor(house$floors, levels =sort(unique(house$floors)))
house %>%
  st_drop_geometry() %>%
  group_by(floor.factor)%>%
  summarize(price_m = mean(price))%>%
  ggplot(aes(x = floor.factor, y = price_m)) +
  geom_col(position = "dodge")+
  plotTheme() + theme(axis.text.x = element_text(angle = 45, hjust = 1)) #1-2+3, 2.5+3.5
house <- house %>%
  mutate(floor_cat = factor(case_when(
    floors <= 2 | floors == 3 ~ "regular",
    floors %in% c(2.5, 3.5) ~ "irregular"
  )))%>%
  select(-floor.factor)
# check the numbers of each category
table(house$floor_cat)

#reclassify
house <- house%>%
  mutate(grade_dum = ifelse(grade <= 9, "low","high"))

#other category variables
colnames(house)
str(house)
house %>% 
  st_drop_geometry()%>%
  dplyr::select(price, reno_dum, bed_cat, bath_dum,floor_cat,
                water_dum,view_cat,condition_cat, grade_dum) %>%
  gather(Variable, Value, -price) %>% 
   ggplot(aes(Value, price)) +
     geom_bar(position = "dodge", stat = "summary", fun.y = "mean") +
     facet_wrap(~Variable, ncol = 4, scales = "free") +
     labs(title = "Price as a function of categorical variables", y = "Mean_Price") +
     plotTheme() + theme(axis.text.x = element_text(angle = 45, hjust = 1))

# select variables
house <- house %>%
  select("key", "price","year_used","reno_dum",
         "bedrooms", "bed_cat", "bathrooms", "bath_dum",
         "sqft_living", "sqft_lot", "floor_cat",
         "water_dum", "view_cat", "condition_cat", "grade_dum")
```

## Socio-economic Characteristics

```{r acs_vars, results='hide'}
census_api_key("3ec2bee8c227ff3f9df970d0ffbb11ee1384076e", install = TRUE, overwrite = TRUE)

acs_variable_list.2015 <- load_variables(2015, # year
                                         "acs5", # five year ACS estimates
                                         cache = TRUE)

acs_vars <- c("B01003_001E", # total population
              "B01001A_001E", # white alone
              "B01001_003E", # male under 5
              "B01001_004E", # male 5-9
              "B01001_005E", # male 10-14
              "B01001_020E", # male 65-66
              "B01001_021E", # male 67-69
              "B01001_022E", # male 70-74
              "B01001_023E", # male 75-79
              "B01001_024E", # male 80-84
              "B01001_025E", # male over 85
              "B01001_027E", # female under 5
              "B01001_028E", # female 5-9
              "B01001_029E", # female 10-14
              "B01001_044E", # female 65-66
              "B01001_045E", # female 67-69
              "B01001_046E", # female 70-74
              "B01001_047E", # female 75-79
              "B01001_048E", # female 80-84
              "B01001_049E", # female over 85
              "B15003_001E", # educational attainment over 25
              "B15003_022E", # bachelor's degree
              "B19013_001E", # median household income
              "B23025_004E", # employed labor force
              "B23025_003E", # total labor force
              "B17020_002E") # income below poverty level
```

- age structure
- reference: "total dependency ratio and elderly dependency ratio are on an inverse relationship towards ordinary residence price." https://www.scirp.org/journal/paperinformation?paperid=74919

- education level
- reference: "The findings show that higher education does have a positive relationship to the house prices in Sweden."
https://www.diva-portal.org/smash/get/diva2:1346009/FULLTEXT01.pdf

```{r get_acs_2015}
acsTractsSeattle.2015 <- get_acs(geography = "tract",
                             year = 2015, 
                             variables = acs_vars,
                             geometry = TRUE,
                             state = "Washington", 
                             county = "King",
                             output = "wide") %>%
  st_transform(st_crs(seattle)) %>%
  select(GEOID, NAME, all_of(acs_vars)) %>%
  rename(total_pop = B01003_001E,
          white_pop = B01001A_001E,
          edu_bach = B15003_022E,
          edu_attain = B15003_001E,
          median_hh_income = B19013_001E,
          total_labor = B23025_003E,
          employ_labor = B23025_004E,
          poverty = B17020_002E) %>%
  mutate(area = st_area(.)) %>%
  mutate(pop_den = ifelse(as.numeric(area) > 0, total_pop / area, 0),
         white_share = round(ifelse(total_pop > 0, white_pop / total_pop, 0) * 100, digits = 2),
         pop_under14 = B01001_003E + B01001_004E + B01001_005E + B01001_027E +
            B01001_028E + B01001_029E,
         pop_over65 = B01001_020E + B01001_021E + B01001_022E + B01001_023E +
            B01001_024E + B01001_025E + B01001_044E + B01001_045E + B01001_046E +
            B01001_047E + B01001_048E + B01001_049E,
         total_dep = round((pop_under14 + pop_over65) / (total_pop -
                               (pop_under14 + pop_over65)) * 100, digits = 2),
         elder_dep = round(pop_over65 / (total_pop - (pop_under14 + pop_over65)) * 100, digits = 2),
         bach_share = round(ifelse(edu_attain > 0, edu_bach/edu_attain, 0) * 100, digits = 2),
         employ_rate = round(ifelse(total_labor > 0, employ_labor / total_labor, 0) * 100, digits = 2),
         pover_rate = round(ifelse(total_pop > 0, poverty / total_pop, 0) * 100, digits = 2))

acsTractsSeattle.2015 <- acsTractsSeattle.2015 %>%
  mutate(bach_dum = ifelse(bach_share > mean(acsTractsSeattle.2015$bach_share,
                                                        na.rm = TRUE), "above", "below"),
         median_hh_dum = ifelse(median_hh_income > mean(acsTractsSeattle.2015$median_hh_income,
                                                        na.rm = TRUE), "above", "below"),
         employ_dum = ifelse(employ_rate > mean(acsTractsSeattle.2015$employ_rate,
                                                        na.rm = TRUE), "above", "below"),
         pover_dum = ifelse(pover_rate > mean(acsTractsSeattle.2015$pover_rate,
                                                        na.rm = TRUE), "above", "below")) %>%
  select(GEOID, NAME, pop_den, white_share, total_dep, elder_dep, bach_share, bach_dum, median_hh_income,
         median_hh_dum, employ_rate, employ_dum, pover_rate, pover_dum)

house <- house %>%
  st_join(., acsTractsSeattle.2015) %>%
  filter(!is.na(pop_den)) # remove NA, one house may outside the boundary of census tracts

# frequency of dummy variables
table(house$bach_dum)
table(house$median_hh_dum)
table(house$employ_dum)
table(house$pover_dum)

## plot mean house price of bach_dum
ggplot(house %>% st_drop_geometry() %>% group_by(bach_dum) %>%
  summarise(aver_price = mean(price))) +
  geom_col(aes(bach_dum,aver_price))

## plot mean price of median_hh_dum
ggplot(house %>% st_drop_geometry() %>% group_by(median_hh_dum) %>%
  summarise(aver_price = mean(price))) +
  geom_col(aes(median_hh_dum,aver_price))

## plot mean price of employ_dum
ggplot(house %>% st_drop_geometry() %>% group_by(employ_dum) %>%
  summarise(aver_price = mean(price))) +
  geom_col(aes(employ_dum,aver_price))

## plot mean price of pover_dum
ggplot(house %>% st_drop_geometry() %>% group_by(pover_dum) %>%
  summarise(aver_price = mean(price))) +
  geom_col(aes(pover_dum,aver_price))

## plot all the mean of dummy variables
house %>% 
  st_drop_geometry()%>%
  dplyr::select(price, bach_dum, median_hh_dum, employ_dum, pover_dum) %>%
  gather(Variable, Value, -price) %>% 
   ggplot(aes(Value, price)) +
     geom_bar(position = "dodge", stat = "summary", fun.y = "mean") +
     facet_wrap(~Variable, ncol = 2, scales = "free") +
     labs(title = "Price as a function of dummy variables", y = "Mean_Price") +
     plotTheme() + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Amenities Characteristics

### subway station

- variable  
1) distance to the nearest station
2) categories based on distance (0-0.5/0.5-1/1+ mile) (0.5mile = 2640ft)  

- reference: "houses within a quarter mile to a half mile of a Metro station sold for 7.5% more, and houses within a half mile to a mile sold for 3.9% more." https://www.freddiemac.com/research/insight/20191002-metro-station-impact#:~:text=Similarly%2C%20houses%20within%20a%20quarter,mile%20sold%20for%203.9%25%20more.

- source: https://gis-kingcounty.opendata.arcgis.com/datasets/7fb1b64925db450e8f024940f697823e_390/explore?location=47.584121%2C-122.115870%2C10.45

```{r subway}
#read station point data
sub <- st_read(here::here("data/raw/Amenities/Metro_Sub_Stations_in_King_County___sub_stations_point.geojson"))%>%
  st_transform(st_crs(house))

# the distance to the nearest station 
house <-  house %>%
      mutate(sub_dis = nn_function(st_coordinates(house),
                                      st_coordinates(sub), k = 1))

# categories based on distance
house <- house %>%
  mutate(sub_cat = case_when(
    sub_dis <= 2640 ~ "within0.5mile",
    sub_dis > 2640 & sub_dis <= 5280 ~ "0.5-1mile",
    sub_dis > 5280 ~ "1+mile"
  ))
```

### school district

- variable: the name of the school district

- reference: "the price differentials for similar homes — same square footage, number of bedrooms and baths — that are located near each other but served by different school districts can range from tens of thousands to hundreds of thousands of dollars." https://www.seattletimes.com/business/how-much-do-good-schools-boost-home-prices/  

- source: https://gis-kingcounty.opendata.arcgis.com/datasets/94eb521c71f2401586c6ce6a34d68166_406/explore?location=47.672575%2C-122.604064%2C18.86

```{r school}
sch <- st_read(here::here("data/raw/Amenities/Seattle_School_Board_Director_Districts___dirdst_area.geojson"))%>%
  st_transform(st_crs(house))

# add the school district variable
house <- house %>% 
  st_join(sch%>%select(DIRDST), join = st_within)%>%
  rename(sch_cat = "DIRDST")
```

### parks

- variable: 
1) area of parks within 500 feet radius
2) count of parks within 500 feet radius

- reference: "Most community-sized parks (~40 acres) had a substantial impact on home prices up to a distance of 500 to 600 feet. While the influence of larger parks extended out to 2,000 feet, beyond 500 feet the influence was relatively small." https://www.naturequant.com/blog/Impact-of-Parks-on-Property-Value/  

- source: https://gis-kingcounty.opendata.arcgis.com/datasets/a0c94c33228146c5ad95a1dff3b6963d_228/explore?location=47.557674%2C-122.213839%2C11.45

```{r park}
park<-st_read(here::here("data/raw/Amenities/Parks_in_King_County___park_area.geojson"))%>%
  st_transform(st_crs(house))%>%
  st_intersection(seattle)%>%
  mutate(park_area = st_area(.))
  
# area and count of parks within 500 feet
house_parks <- st_join(st_buffer(house, dist = 500), park, join = st_intersects)%>%
  group_by(key) %>% 
  summarise(park_c = n_distinct(SITENAME, na.rm = TRUE),
            sum_park_area = sum(park_area, na.rm = TRUE))
house_parks$all_park_area <- as.numeric(house_parks$sum_park_area)
house_parks$park_cat <- as.factor(house_parks$park_c)

house <- house %>% 
    left_join(house_parks%>%
                st_drop_geometry()%>%
                select(key,park_cat, all_park_area), by = "key")%>%
  rename(parks_area = all_park_area)
```

### tree canopy

- variable:
1) percent of existing tree canopy (2016, hexagons)

- reference: "a city facing major development pressure — trees were associated with an increase in single family home values"
https://www.vibrantcitieslab.com/resources/urban-trees-increase-home-values/

- source: https://data-seattlecitygis.opendata.arcgis.com/datasets/SeattleCityGIS::seattle-tree-canopy-2016-2021-50-acre-hexagons/explore?layer=1&location=47.580733%2C-122.309741%2C11.26

```{r tree_canopy}
tree_canopy_2016 <- st_read(here::here("data/raw/Amenities/Tree_Canopy.geojson"))%>%
  st_transform(st_crs(house)) %>%
  mutate(tree_canopy = round(TreeCanopy_2016_Percent, digits = 2)) %>%
  select(tree_canopy)

house <- house %>%
  st_join(., tree_canopy_2016)
```

### medical facilities

- variable: 
1) average distance to the nearest 1/2/3 medical facilities
2) categories based on distance (0-0.5/0.5-1/1+ mile)

- reference: "hospitals would only be highly evaluated in a ‘close-but-not-too-close’ geographic location" https://www.researchgate.net/publication/282942128_The_non-linearity_of_hospitals'_proximity_on_property_prices_experiences_from_Taipei_Taiwan

- source: https://gis-kingcounty.opendata.arcgis.com/datasets/1b7f0fb5179a400f91a35c0b6bfd77c9_733/explore

```{r medical}
med <- st_read(here::here("data/raw/Amenities/Medical_Facilities_including_Hospitals___medical_facilities_point.geojson"))%>%
  st_transform(st_crs(house))%>%
  st_intersection(seattle)

# calculate the distance to the nearest medical facilities 
house <-  house %>%
      mutate(med_dis1 = nn_function(st_coordinates(house),
                                      st_coordinates(med), k = 1),
             med_dis2 = nn_function(st_coordinates(house),
                                      st_coordinates(med), k = 2),
             med_dis3 = nn_function(st_coordinates(house),
                                      st_coordinates(med), k = 3))

# categories based on distance
house <- house %>%
  mutate(med_cat = case_when(
    med_dis1 <= 2640 ~ "within0.5mile",
    med_dis1 > 2640 & med_dis1 <= 5280 ~ "0.5-1mile",
    med_dis1 > 5280 ~ "1+mile"
  ))
```

### commercial

- variable:  
1) average distance to the nearest 1/2/3 shops
2) categories based on distance (0-0.5/0.5-1/1+ mile)

- reference: 
1) "an inverse relationship between the housing price and its distance from the shopping mall" https://www.diva-portal.org/smash/get/diva2:1450713/FULLTEXT01.pdf  
2) "Notwithstanding the negative externalities of shopping centres, residential
properties within 100-metre radius of shopping centres command a higher premium
than those farther away although the price-distance relationship is not monotonic
while the proximity factor varies from housing estate to housing estate." https://www.prres.org/uploads/746/1752/Addae-Dapaah_Shopping_Centres_Proximate_Residential_Properties.pdf

- source:https://gis-kingcounty.opendata.arcgis.com/datasets/4fdb4709874b46cf8dbb284182ca0094_383/explore?showTable=true  
Select type by "CODE": https://www.arcgis.com/sharing/rest/content/items/4fdb4709874b46cf8dbb284182ca0094/info/metadata/metadata.xml?format=default&output=html
- 690 Shopping Centers

```{r commercial}
mark<- st_read(here::here("data/raw/Amenities/King_County_Landmarks___landmark_point.geojson"))%>%
  st_transform(st_crs(house))%>%
  st_intersection(seattle)

#select shopping center
mark_shop <- mark%>%
  filter(CODE == 690)

# calculate the distance to the nearest 1/2/3 shopping center(s) 
house <- house %>%
      mutate(shop_dis1 = nn_function(st_coordinates(house),
                                      st_coordinates(mark_shop), k = 1),
             shop_dis2 = nn_function(st_coordinates(house),
                                      st_coordinates(mark_shop), k = 2),
             shop_dis3 = nn_function(st_coordinates(house),
                                      st_coordinates(mark_shop), k = 3))

# categories based on distance
house <- house %>%
  mutate(shop_cat = case_when(
    shop_dis1 <= 2640 ~ "within0.5mile",
    shop_dis1 > 2640 & shop_dis1 <= 5280 ~ "0.5-1mile",
    shop_dis1 > 5280 ~ "1+mile"
  ))
```

### crime

- variable:
1) count of crimes within a 1/8 mi
2) average distance to the nearest 1/2/3/4/5 crimes (robbery and  assault only)

- reference:"only robbery and aggravated assault crimes (per acre) exert a meaningful influence upon neighborhood housing values." https://sciencedirect.com/science/article/pii/S0166046210000086#aep-abstract-id3
"The overall effect on house prices of crime (measured as crime rates) is relatively small, but if its impact is measured by distance to a crime hot spot, the effect is non-negligible." https://www.researchgate.net/publication/335773241_Do_crime_hot_spots_affect_housing_prices_Do_crime_hot_spots_affect_housing_prices

- source:https://data.seattle.gov/Public-Safety/SPD-Crime-Data-2008-Present/tazs-3rd5/about_data

**To avoid rerunning the huge dataset:**
```{r crime_clean, eval}
# crime <- read.csv(here::here("data/raw/Amenities/SPD_Crime_Data__2008-Present_20240328.csv"))

# get the target crime
# crime_clean <- crime %>%
#   mutate(year = str_sub(Report.Number, 1, 4)) %>%
#   filter(year %in% c("2013", "2014", "2015") )%>% #choose those before and in 2015
#   filter(!Offense %in% c(
#     "Bad Checks",
#     "Bribery",
#     "Embezzlement",
#     "Extortion/Blackmail",
#     "Credit Card/Automated Teller Machine Fraud",
#     "False Pretenses/Swindle/Confidence Game",
#     "Identity Theft",
#     "Impersonation",
#     "Welfare Fraud",
#     "Wire Fraud",
#     "Curfew/Loitering/Vagrancy Violations",
#     "Driving Under the Influence",
#     "Drug Equipment Violations",
#     "Drug/Narcotic Violations",
#     "Betting/Wagering",
#     "Gambling Equipment Violation",
#     "Operating/Promoting/Assisting Gambling",
#     "Liquor Law Violations",
#     "Pornography/Obscene Material",
#     "Assisting or Promoting Prostitution",
#     "Prostitution",
#     "Weapon Law Violations"
#   ))%>% #exclude those with little impact on housing price e.g.Financial Crimes, Public Order Offenses
#   filter(Longitude != 0 & Latitude != 0)# select the valid ones
#
# write.csv(crime_clean , here::here("data/processed/crime_clean.csv"), row.names = FALSE)

# read the cleaned data set
crime <- read.csv(here::here("data/processed/crime_clean.csv")) %>%
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326)%>%
  st_transform(st_crs(house))

# count of crime within 1/8 mi
house$crime_c <- house %>%
    st_buffer(660) %>%
    aggregate(mutate(crime, counter = 1)%>%select(counter),., sum) %>%
    pull(counter)

# calculate the distance to the nearest 1/2/3/4/5 crime locations
house <-  house %>%
      mutate(crime_dis1 = nn_function(st_coordinates(house),
                                      st_coordinates(crime), k = 1),
             crime_dis2 = nn_function(st_coordinates(house),
                                      st_coordinates(crime), k = 2),
             crime_dis3 = nn_function(st_coordinates(house),
                                      st_coordinates(crime), k = 3),
             crime_dis4 = nn_function(st_coordinates(house),
                                      st_coordinates(crime), k = 4),
             crime_dis5 = nn_function(st_coordinates(house),
                                      st_coordinates(crime), k = 5))
```

### inspect categories

```{r amenities category}
#plot all the mean of categorical variables
house %>% 
  st_drop_geometry()%>%
  dplyr::select(price ,sub_cat, park_cat, med_cat,shop_cat,sch_cat) %>%
  gather(Variable, Value, -price) %>% 
   ggplot(aes(Value, price)) +
     geom_bar(position = "dodge", stat = "summary", fun.y = "mean") +
     facet_wrap(~Variable, ncol = 2, scales = "free") +
     labs(title = "Price as a function of categorical variables", y = "Mean_Price") +
     plotTheme() + theme(axis.text.x = element_text(angle = 45, hjust = 1))
#-> med_cat not useful

house <- house %>%
  select(-med_cat)
```

## Spatial Structure

```{r fixed_effect}
# large neighborhoods
neigh_large <- st_read(here::here("data/raw/Boundary/Neighborhood_Map_Atlas_Districts.geojson")) %>%
  st_transform(st_crs(house)) %>%
  rename(L_NAME = L_HOOD) %>%
  select(L_NAME)

# small neighborhoods
neigh_small <- st_read(here::here("data/raw/Boundary/Neighborhood_Map_Atlas_Neighborhoods.geojson")) %>%
  st_transform(st_crs(house)) %>%
  rename(S_NAME = S_HOOD) %>%
  select(S_NAME)

# census tracts
neigh_tract <- acsTractsSeattle.2015 %>%
  select(NAME) %>%
  rename(T_NAME = NAME) %>%
  st_intersection(seattle)
```


# Data Description

## Summary Table

```{r continuous_summary}
# continuous variables
continuous_summary <- house %>%
  st_drop_geometry() %>%
  select(- key)%>%
  summarise(across(where(is.numeric),
                   list(maximum = ~ max(., na.rm = TRUE),
                        minimum = ~ min(., na.rm = TRUE),
                        mean = ~ mean(., na.rm = TRUE),
                        standard_deviation = ~ sd(., na.rm = TRUE),
                        n = ~ sum(!is.na(.x))),
                   .names = "{.col}:{.fn}")) %>%
  pivot_longer(cols = everything(), names_to = c("variables", "statistic"),
               names_sep = ":", values_to = "value") %>%
  mutate(value = round(value, digits = 2)) %>%
  pivot_wider(names_from = statistic, values_from = value)

house_continuous <- house %>%
  st_drop_geometry() %>%
  select(- key) %>%
  select_if(is.numeric)%>%
  na.omit()
  

stargazer(house_continuous,
          type = "text",
          digits = 1,
          title = "Summary Statistics")

continuous_description <- data.frame(
  variables = continuous_summary$variables,
  category = rep(c("independent","internal", "socio-economic", "amenities"),
                 times = c(1, 5, 8, 15)),
  description = c("price of each unit",
                  "years from built to 2015",
                  "the number of bedrooms in each unit",
                  "the number of bathrooms in each unit",
                  "the area of living of each unit",
                  "the area of parking lot of each unit",
                  "the number of population per square feet in the census tract",
                  "the ratio of white people to total population in the census tract",
                  "the ratio of the number of children (0-14 years old) and older persons (65 years or over) to the working-age population (15-64 years old) in the census tract",
                  "the ratio of older persons (65 years or over) to the working-age population (15-64 years old) in the census tract",
                  "the percentage of with a bachelor's degree among adults age 25 and older in the census tract",
                  "median househhold income in the census tract",
                  "the ratio of the employed to the working age population in the census tract",
                  "the ratio of the number of people (in a given age group) whose income falls below the poverty line to total population in the census tract",
                  "the distance to the nearest subway station",
                  "the total area of parks located within a 500-foot radius of each unit",
                  "the ratio of the area of tree canopy to the total area in the measuring space",
                  "the distance to the nearest medical facility",
                  "the average distance to the nearest 2 medical facilities",
                  "the average distance to the nearest 3 medical facilities",
                  "the distance to the nearest shopping center",
                  "the average distance to the nearest 2 shopping center",
                  "the average distance to the nearest 3 shopping center",
                  "the number of crimes within a 1/8-mile radius around each unit",
                  "the distance to the nearest crime",
                  "the average distance to the nearest 2 crime",
                  "the average distance to the nearest 3 crime",
                  "the average distance to the nearest 4 crime",
                  "the average distance to the nearest 5 crime"),
  unit = c("$",
           "year", "-","-","sqft","sqft",
           "person / sqft", "%","%","%","%","$","%","%",
           "feet","sqft","%","feet","feet","feet","feet","feet","feet",
           "-", "feet","feet","feet","feet","feet"
           ))
continuous_description %>%
  left_join(continuous_summary, by = "variables")%>%
  kable() %>%
  kable_styling()
```
```{r dummy summary}
str(house[is.character()])
# dummy variables
dummy_summary <- house %>%
  st_drop_geometry() %>%
  summarise(across(c(reno_dum, bath_dum,water_dum,grade_dum)),
            group_by(.) %>%
              summarise(count = n()) %>%
              mutate(percent = round(count/sum(count) * 100, digits = 2)))

dummy_summary <- house %>%
  st_drop_geometry() %>%
  select(reno_dum, bath_dum, water_dum, grade_dum) %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value") %>%
  group_by(variable, value) %>%
  summarise(count = n(), .groups = 'drop') %>%
  mutate(percent = round((count / sum(count)) * 100, digits = 2)) %>%
  ungroup()





```



# Exploratory Data Analysis

```{r correlation}
house %>%
  st_drop_geometry() %>%
  pivot_longer(cols = where(is.numeric), names_to = "variable", values_to = "value")%>%
  ggplot(aes(x = value)) +
  geom_histogram(bins = 30, fill = "blue", color = "black") + # adjust bin count as needed
  facet_wrap(~variable,ncol = 4,  scales = "free_x") + # separate plot for each variable
  labs(x = "Value", y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
  

```


## Mapping

```{r price_map}
palette5 <- brewer.pal(5, "YlGnBu")

# ggplot, reorder
boston.sf$quintiles <- ntile(boston.sf$PricePerSq, 5)

# mapping data
ggplot() +
  geom_sf(data = nhoods, fill = "lightgrey", col = "white") +
  geom_sf(data = boston.sf, aes(colour = q5(PricePerSq)), 
          show.legend = "point", size = .75) +
  scale_colour_manual(values = palette5,
                   labels=qBr(boston,"PricePerSq"),
                   name="Quintile\nBreaks") +
  labs(title="Price Per Square Foot, Boston") +
  theme_void()
```

## Four Home Price Correlation Scatterplots

### White Population Share

```{r white_share}
ggplot(house) +
  geom_point(aes(x = log(job_density + 1), y = log(taxi_trip_2017)), color = "black", pch = 16, size = 2) +
  labs(title = "2017 Taxi Trips vs. Job Density",
       x = "Employment per Acre (logged)",
       y = "Taxi Trips (logged)") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
```

# Modeling

Process:
1) whole data set -> run lm -> select one var in each category
2) split data set -> run lm on trainning data set -> run vif and cor -> deal with multicolinearity
3) delete insignificant var. until all var at p < 0.05
4) delete var. with large p and examine R^2 & residual

## Original House Price

## Initial Variables Selection


```{r selection}
# finalize regression dataset
house <- house %>%
  mutate(parks_area_log = log(parks_area + 1)) %>%
  select(-GEOID, -NAME) %>%
  st_drop_geometry()
str(house)

# select continuous or categorical variable
lm1 <- lm(price ~ .-key, data = house)
summary(lm1)

## bedroom
lm2 <- lm(price ~ .-bed_cat-key, data = house)
summary(lm2)

house <- house %>%
  select(-bed_cat)

## bathroom
lm3 <- lm(price ~ .-bathrooms-key, data = house)
summary(lm3)

house <- house %>%
  select(-bathrooms)

## dependency
lm4 <- lm(price ~ .-total_dep-key, data = house)
summary(lm4)

house <- house %>%
  select(-total_dep)

## bachelor degree
lm5 <- lm(price ~ .-bach_dum-key, data = house)
summary(lm5)

house <- house %>%
  select(-bach_dum)

## median household income
lm6 <- lm(price ~ .-median_hh_dum-key, data = house)
summary(lm6)

house <- house %>%
  select(-median_hh_dum)

## employment
lm7 <- lm(price ~ .-employ_dum-key, data = house)
summary(lm7)

house <- house %>%
  select(-employ_dum)

## poverty
lm8 <- lm(price ~ .-pover_dum-key, data = house)
summary(lm8)

house <- house %>%
  select(-pover_dum)

## subway station
lm9 <- lm(price ~ .-sub_cat-key, data = house)
summary(lm9)

house <- house %>%
  select(-sub_cat)

## park
lm10 <- lm(price ~ .-parks_area-parks_area_log-key, data = house)
summary(lm10)

house <- house %>%
  select(-parks_area, -parks_area_log)

## medical facility
lm11 <- lm(price ~ .-med_dis2-med_dis3-key, data = house)
summary(lm11)

house <- house %>%
  select(-med_dis2, -med_dis3)

## shopping center
lm12 <- lm(price ~ .-shop_dis2-shop_dis3-key, data = house)
summary(lm12)

house <- house %>%
  select(-shop_dis2, -shop_dis3)

lm13 <- lm(price ~ .-shop_cat-key, data = house)
summary(lm13)

house <- house %>%
  select(-shop_cat)

## crime
lm14 <- lm(price ~ .-crime_dis2-crime_dis3-crime_dis4-crime_dis5-key, data = house)
summary(lm14)

house <- house %>%
  select(-crime_dis2, -crime_dis3, -crime_dis4, -crime_dis5)

lm15 <- lm(price ~ .-crime_dis1-key, data = house)
summary(lm15)

house <- house %>%
  select(-crime_dis1)

## add fixed effect
house_lm <- house %>%
  left_join(hh%>%select(key), by = "key") %>%
  st_as_sf() %>%
  st_join(., neigh_large) %>%
  st_join(., neigh_small) %>%
  st_join(., neigh_tract) %>%
  st_drop_geometry() %>%
  select(-key)
str(house_lm)
```

## Split Training and Testing Datasets

```{r traintest}
# check frequency of categorical variables
table(house_lm$reno_dum)
table(house_lm$bath_dum)
table(house_lm$floor_cat)
table(house_lm$water_dum)
table(house_lm$view_cat)
table(house_lm$condition_cat)
table(house_lm$grade_dum)
table(house_lm$sch_cat)
table(house_lm$park_cat)
table(house_lm$L_NAME)

# split data 0.7/0.3
set.seed(1)
inTrain <- createDataPartition(
              y = paste(house_lm$reno_dum, house_lm$bath_dum, house_lm$floor_cat,
                        house_lm$water_dum, house_lm$view_cat, house_lm$condition_cat,
                        house_lm$grade_dum, house_lm$sch_cat, house_lm$park_cat, house_lm$L_NAME), 
              p = .70, list = FALSE)  # create a vector for the training set (70%)

# subset the dataset to create the training set
seattle.train.lm <- house_lm[inTrain,] # training set

# subset the dataset to create the testing set
seattle.test.lm <- house_lm[-inTrain,] # testing set
```

## Build Multiple Linear Regression Model

```{r multiple_regression}
# ignore neighborhood effect first
seattle.train <- seattle.train.lm %>%
  select(-L_NAME, -S_NAME, -T_NAME)

seattle.test <- seattle.test.lm %>%
  select(-L_NAME, -S_NAME, -T_NAME)

# build regression model on training dataset
lm16 <- lm(price ~ ., data = seattle.train)
summary(lm16)
vif(lm16)
str(seattle.train)
cor(seattle.train%>%select(-price, -reno_dum, -bath_dum, -floor_cat, -water_dum,
                           -view_cat, -condition_cat, -grade_dum, -sch_cat, -park_cat))

## sub_dis
lm17 <- lm(sub_dis ~ reno_dum + bath_dum + floor_cat + water_dum + view_cat +
             condition_cat + grade_dum + sch_cat + park_cat, data = seattle.train)
summary(lm17)
vif(lm17)

lm18 <- lm(price ~ .-sub_dis, data = seattle.train)
summary(lm18)
vif(lm18)

seattle.train <- seattle.train %>%
  select(-sub_dis)

## high correlation: white share & bachelor degree
lm19 <- lm(price ~ .-bach_share, data = seattle.train)
summary(lm19)
vif(lm19)

seattle.train <- seattle.train %>%
  select(-bach_share)
cor(seattle.train%>%select(-price, -reno_dum, -bath_dum, -floor_cat, -water_dum,
                           -view_cat, -condition_cat, -grade_dum, -sch_cat, -park_cat))

## high correlation: white share & poverty rate
lm20 <- lm(price ~ .-pover_rate, data = seattle.train)
summary(lm20)
vif(lm20)

seattle.train <- seattle.train %>%
  select(-pover_rate)
cor(seattle.train%>%select(-price, -reno_dum, -bath_dum, -floor_cat, -water_dum,
                           -view_cat, -condition_cat, -grade_dum, -sch_cat, -park_cat))

## insignificance: tree canopy
lm21 <- lm(price ~ .-tree_canopy, data = seattle.train)
summary(lm21)
vif(lm21)

seattle.train <- seattle.train %>%
  select(-tree_canopy)
```

## Test on testing dataset

``` {r modify_model}
# make predictions on the test set and evaluate model performance
lm22 <- lm(price ~ ., data = seattle.train)
summary(lm22) # adjusted R-squared = 0.802
plot(lm22)

## first regression model from last step
seattle.test1 <-
  seattle.test %>%
  mutate(regression = "baseline regression", # add a column indicating the type of regression model used
         price.predict = predict(lm22, seattle.test), # predict house prices using the trained regression model
         # calculate the difference between predicted and actual house prices
         price.error = price.predict - price,
         # calculate the absolute difference between predicted and actual house prices
         price.AbsError = abs(price.predict - price),
         # calculate the absolute percentage error
         price.APE = round((abs(price.predict - price)) / price * 100, digits = 2))

mean(seattle.test1$price.AbsError) # 94490.7
mean(seattle.test1$price.APE) # 19.10%

## remove "year_used"
lm23 <- lm(price ~ .-year_used, data = seattle.train)
summary(lm23) # adjusted R-squared = 0.8019

seattle.test2 <-
  seattle.test %>%
  mutate(regression = "baseline regression", # add a column indicating the type of regression model used
         price.predict = predict(lm23, seattle.test), # predict house prices using the trained regression model
         # calculate the difference between predicted and actual house prices
         price.error = price.predict - price,
         # calculate the absolute difference between predicted and actual house prices
         price.AbsError = abs(price.predict - price),
         # calculate the absolute percentage error
         price.APE = round((abs(price.predict - price)) / price * 100, digits = 2))

mean(seattle.test2$price.AbsError) # 94371.65
mean(seattle.test2$price.APE) # 19.05%

seattle.train <- seattle.train %>%
  select(-year_used)

## remove "crime_c"
lm24 <- lm(price ~ .-crime_c, data = seattle.train)
summary(lm24) # adjusted R-squared = 0.8015

seattle.test3 <-
  seattle.test %>%
  mutate(regression = "baseline regression", # add a column indicating the type of regression model used
         price.predict = predict(lm24, seattle.test), # predict house prices using the trained regression model
         # calculate the difference between predicted and actual house prices
         price.error = price.predict - price,
         # calculate the absolute difference between predicted and actual house prices
         price.AbsError = abs(price.predict - price),
         # calculate the absolute percentage error
         price.APE = round((abs(price.predict - price)) / price * 100, digits = 2))

mean(seattle.test3$price.AbsError) # 94189.17
mean(seattle.test3$price.APE) # 19.00%

seattle.train <- seattle.train %>%
  select(-crime_c)

## remove "employ_rate"
lm25 <- lm(price ~ .-employ_rate, data = seattle.train)
summary(lm25) # adjusted R-squared = 0.8005

seattle.test4 <-
  seattle.test %>%
  mutate(regression = "baseline regression", # add a column indicating the type of regression model used
         price.predict = predict(lm25, seattle.test), # predict house prices using the trained regression model
         # calculate the difference between predicted and actual house prices
         price.error = price.predict - price,
         # calculate the absolute difference between predicted and actual house prices
         price.AbsError = abs(price.predict - price),
         # calculate the absolute percentage error
         price.APE = round((abs(price.predict - price)) / price * 100, digits = 2))

mean(seattle.test4$price.AbsError) # 93891.87
mean(seattle.test4$price.APE) # 18.93%

seattle.train <- seattle.train %>%
  select(-employ_rate)

## remove "med_dis1"
lm26 <- lm(price ~ .-med_dis1, data = seattle.train)
summary(lm26) # adjusted R-squared = 0.7995

seattle.test5 <-
  seattle.test %>%
  mutate(regression = "baseline regression", # add a column indicating the type of regression model used
         price.predict = predict(lm26, seattle.test), # predict house prices using the trained regression model
         # calculate the difference between predicted and actual house prices
         price.error = price.predict - price,
         # calculate the absolute difference between predicted and actual house prices
         price.AbsError = abs(price.predict - price),
         # calculate the absolute percentage error
         price.APE = round((abs(price.predict - price)) / price * 100, digits = 2))

mean(seattle.test5$price.AbsError) # 93529.29
mean(seattle.test5$price.APE) # 18.79%

seattle.train <- seattle.train %>%
  select(-med_dis1)

## remove "elder_dep"
lm27 <- lm(price ~ .-elder_dep, data = seattle.train)
summary(lm27) # adjusted R-squared = 0.7965

seattle.test6 <-
  seattle.test %>%
  mutate(regression = "baseline regression", # add a column indicating the type of regression model used
         price.predict = predict(lm27, seattle.test), # predict house prices using the trained regression model
         # calculate the difference between predicted and actual house prices
         price.error = price.predict - price,
         # calculate the absolute difference between predicted and actual house prices
         price.AbsError = abs(price.predict - price),
         # calculate the absolute percentage error
         price.APE = round((abs(price.predict - price)) / price * 100, digits = 2))

mean(seattle.test6$price.AbsError) # 93115.74
mean(seattle.test6$price.APE) # 18.59%

seattle.train <- seattle.train %>%
  select(-elder_dep)

## remove "pop_den"
lm28 <- lm(price ~ .-pop_den, data = seattle.train)
summary(lm28) # adjusted R-squared = 0.7965

seattle.test7 <-
  seattle.test %>%
  mutate(regression = "baseline regression", # add a column indicating the type of regression model used
         price.predict = predict(lm28, seattle.test), # predict house prices using the trained regression model
         # calculate the difference between predicted and actual house prices
         price.error = price.predict - price,
         # calculate the absolute difference between predicted and actual house prices
         price.AbsError = abs(price.predict - price),
         # calculate the absolute percentage error
         price.APE = round((abs(price.predict - price)) / price * 100, digits = 2))

mean(seattle.test7$price.AbsError) # 93073.76
mean(seattle.test7$price.APE) # 18.57%

seattle.train <- seattle.train %>%
  select(-pop_den)
```

## Cross Validation

```{r cv}
# without fixed effect
set.seed(1)
fitControl <- trainControl(method = "cv", number = 100)

seattle.cv <- 
  train(price ~ ., data = house_lm %>% select(price, reno_dum, bedrooms, bath_dum,
                                              sqft_living, sqft_lot, floor_cat,
                                              water_dum, view_cat, condition_cat,
                                              grade_dum, white_share, median_hh_income,
                                              sch_cat, park_cat, shop_dis1),
     method = "lm", trControl = fitControl, na.action = na.pass)

seattle.cv

# with fixed effect
set.seed(1)
fitControl <- trainControl(method = "cv", number = 100)

seattle.neigh.cv <- 
  train(price ~ ., data = house_lm %>% select(price, reno_dum, bedrooms, bath_dum,
                                              sqft_living, sqft_lot, floor_cat,
                                              water_dum, view_cat, condition_cat,
                                              grade_dum, white_share, median_hh_income,
                                              sch_cat, park_cat, shop_dis1, T_NAME),
     method = "lm", trControl = fitControl, na.action = na.pass)

seattle.neigh.cv
```





# Conclusion


# Plot Reference

```{r}
# plot Seattle house location
ggplot() + 
    geom_sf(data = neigh_large, color = "red", linewidth = 0.5) +  # Add Seattle boundary
    geom_sf(data = house, color = "black", size = 0.1, pch = 16) +
  #geom_sf(data = crime, color = "green4", size = 0.01, pch = 16)+
    #geom_sf(data = mark_commer, color = "blue", size = 0.3, pch = 16) + 
    #geom_sf(data = poi_commer, color = "red", size = 0.3, pch = 16) + 
    labs(title = "Seattle House Location") +  # Set plot title
    theme_void() +
    theme(plot.title = element_text(hjust = 0.5, face = "bold"))
```


